Provisional Title

### 2021-04-29 draft/draft.tex
changes to fig1, adjustments in analyzeData.R for easier to streamline plots

% packages
\usepackage[utf8]{inputenc} % allows usage of spanish special characters
\usepackage[spanish,english]{babel} % english dictionary for proper hyphenation
\usepackage{amsmath} % math expressions
\usepackage{upgreek} % upright greek letters for math
\usepackage{txfonts} % nice fonts 
\usepackage{authblk} % author affiliations
\usepackage{graphicx} % images
\usepackage{float} % image positioning
\usepackage{multirow} % allows merging cells in tables
\usepackage{makecell} % more table customization
\usepackage{mathtools} % nice matrices
\usepackage{units} % nice fractions
\usepackage{rotating} % rotated text
\usepackage{caption} % customization of figure/table captions
\usepackage{subcaption} % collages of multiple images
\usepackage{hyperref} % hyperlinks
\usepackage{nameref} % cross-references between draft and supplementary material
\usepackage{zref-xr,zref-user} % more cross-references formatting
\usepackage[square,numbers,sort&compress]{natbib}
\usepackage[table,xcdraw]{xcolor} % text and table background colors
\usepackage[margin=3cm]{geometry} % margins
\usepackage[nottoc,numbib]{tocbibind} % add references to table of contents
\usepackage{lineno} % line numbers

@@

Communities were stabilized \textit{ex situ} as described in \cite{Goldford2018}.
In short, environmental samples (soil, leaves...) within one meter radius in eight different
geographical locations were collected with sterile
tweezers or spatulas into 50mL sterile tubes (Fig. \mr).
One gram of each sample was allowed to
sit at room temperature in 10mL of phosphate buffered saline (1$\times$PBS) containing
200$\upmu$g/mL cycloheximide to suppress eukaryotic growth.
After 48h, samples were mixed 1:1 with 80\% glycerol and kept frozen at -80$^\circ$C.
Starting microbial communities were prepared by scrapping the frozen stocks into
200$\upmu$L of 1$\times$PBS and adding a volume of 4$\upmu$L to 500$\upmu$L
of synthetic minimal media (1$\times$M9) supplemented with 200$\upmu$g/mL cycloheximide
and 0.07 C-mol/L glutamine or sodium citrate as the carbon source in 96 deep-well plates
(1.2mL; VWR).
Cultures were then incubated still at 30$^\circ$C to allow for re-growth.
After 48h, samples were fully homogenized and biomass increase was followed by measuring
the optical density (620nm) of 100$\upmu$L of the cultures in a Multiskan FC plate reader
(Thermo Scientific).
Communities were stabilized \cite{Goldford2018} by passaging 4$\upmu$L of the cultures into
500$\upmu$L of fresh media (1$\times$M9 with the carbon source)  every 48h for a total of
12 transfers at a dilution factor of 1:100,
roughly equivalent to 80 generations per culture (Fig. \mr).
Cycloheximide was not added to the media after the first two transfers.

@@

For each community, the most abundant colony morphotype at the end of the ninth transfer
was selected, resuspended in 100$\upmu$L 1$\times$PBS and serially diluted (1:10).
Next, 20$\upmu$L of the cells diluted to $10^{-6}$ were plated in the corresponding synthetic
minimal media and allowed to regrow at 30$^\circ$C for 48h. Dominants were then inoculated
into 500$\upmu$L of fresh media and incubated still at 30$^\circ$C for 48h.
After this period, the communities stabilized for eleven transfers and the isolated dominants
were ready for the competition experiments (Fig \mr) at the onset of the twelfth transfer.

@@

All possible pairwise dominant-dominant and community-community
competition experiments
were performed by mixing equal volumes (4$\upmu$L) of each of the eight
communities or eight dominants at the onset of the twelfth transfer.
Competitions were set up in their native media,
i.e. in 500$\upmu$L of 1$\times$M9 supplemented with 0.07 C-mol/L of
either glutamine or citrate in 96 deep-well plates.
Plates were incubated at 30$^\circ$C for 48h.
Pairwise competitions were further propagated for seven serial transfers
(roughly 42 generations; Fig. \mr) by transferring 8$\upmu$L of
each culture to fresh media (500$\upmu$L).

@@

The sequencing protocol was identical to that described in \cite{Goldford2018}.
Community samples were collected by spinning down at 3500rpm for 25min
in a bench-top centrifuge at room temperature;
cell pellets were stored at -80$^\circ$C before processing.
To maximize Gram-positive bacteria cell wall lysis,
the cell pellets were re-suspended and incubated at 37$^\circ$C for 30min
in enzymatic lysis buffer (20mM Tris-HCl, 2mM sodium EDTA, 1.2\% Triton X-100)
and 20 mg/mL of lysozyme from chicken egg white (Sigma-Aldrich).
After cell lysis, the DNA extraction and purification was performed using the
DNeasy 96 protocol for animal tissues (Qiagen).
The clean DNA in 100$\upmu$L elution buffer of 10mM Tris-HCl, 0.5mM EDTA
at pH 9.0 was quantified using Quan-iT PicoGreen dsDNA Assay Kit
(Molecular Probes, Inc.)
and normalized to 5ng/$\upmu$L in nuclease-free water (Qiagen)
for subsequent 16S rRNA illumina sequencing.
16S rRNA amplicon library preparation was performed following a dual-index
paired-end approach \cite{Kozich2013}.
Briefly, PCR amplicon libraries of V4 regions of the 16S rRNA were prepared 
sing dual-index primers (F515/R805), then pooled and sequenced
using the Illumina MiSeq chemistry and platform.
Each sample went through a 30-cycle PCR in duplicate of 20$\upmu$L
reaction volumes using 5ng of DNA each, dual index primers, and AccuPrime Pfx SuperMix (Invitrogen).
The thermocycling procedure includes a 2min initial denaturation step at
95$^\circ$C, and 30 cycles of the following PCR scheme:
(a) 20 second denaturation at 95$^\circ$C,
(b) 15 second annealing at 55$^\circ$C,
and (c) 5 minute extension at 72$^\circ$C.
The duplicate PCR products of each sample were pooled, purified, and normalized
using SequalPrep PCR cleanup and normalization kit (Invitrogen).
Barcoded amplicon libraries were then pooled and sequenced using
Illumina Miseq v2 reagent kit, which generated 2$\times$250bp paired-end reads
at the Yale Center for Genome Analysis (YCGA).
The sequencing reads were demultiplexed on QIIME 1.9.0 \cite{Caporaso2010}.
The barcodes, indexes, and primers were removed from raw reads,
producing FASTQ files with both the forward and reverse reads for each sample,
ready for DADA2 analysis \cite{Callahan2017}.
DADA2 version 1.1.6 was used to infer unique biological exact sequence variants
(ESVs) for each sample
and na{\"i}ve Bayes was used to assign taxonomy using the SILVA version 123
database \cite{Wang2007,Quast2013}.

>>>>

% packages
\usepackage[utf8]{inputenc} % allows usage of spanish special characters
\usepackage[spanish,english]{babel} % english dictionary for proper hyphenation
\usepackage{amsmath} % math expressions
\usepackage{upgreek} % upright greek letters for math
\usepackage{txfonts} % nice fonts 
\usepackage{authblk} % author affiliations
\usepackage{graphicx} % images
\usepackage{float} % image positioning
\usepackage{multirow} % allows merging cells in tables
\usepackage{makecell} % more table customization
\usepackage{mathtools} % nice matrices
\usepackage{units} % nice fractions
\usepackage{rotating} % rotated text
\usepackage{caption} % customization of figure/table captions
\usepackage{subcaption} % collages of multiple images
\usepackage{hyperref} % hyperlinks
\usepackage{nameref} % cross-references between draft and supplementary material
\usepackage{zref-xr,zref-user} % more cross-references formatting
\usepackage[square,numbers,sort&compress]{natbib}
\usepackage[table,xcdraw]{xcolor} % text and table background colors
\usepackage[margin=3cm]{geometry} % margins
\usepackage{lineno} % line numbers

@@

Communities were stabilized \textit{ex situ} as described in \cite{Goldford2018}.
In short, environmental samples (soil, leaves...) within one meter radius in eight different
geographical locations were collected with sterile
tweezers or spatulas into 50mL sterile tubes (Fig. \mr).
One gram of each sample was allowed to
sit at room temperature in 10mL of phosphate buffered saline (1$\times$PBS) containing
200$\upmu$g/mL cycloheximide to suppress eukaryotic growth.
After 48h, samples were mixed 1:1 with 80\% glycerol and kept frozen at $-80^\circ$C.
Starting microbial communities were prepared by scrapping the frozen stocks into
$200\upmu$L of 1$\times$PBS and adding a volume of $4\upmu$L to $500\upmu$L
of synthetic minimal media (1$\times$M9) supplemented with $200\upmu$g/mL cycloheximide
and 0.07 C-mol/L glutamine or sodium citrate as the carbon source in 96 deep-well plates
(1.2mL; VWR).
Cultures were then incubated still at $30^\circ$C to allow for re-growth.
After 48h, samples were fully homogenized and biomass increase was followed by measuring
the optical density (620nm) of $100\upmu$L of the cultures in a Multiskan FC plate reader
(Thermo Scientific).
Communities were stabilized \cite{Goldford2018} by passaging $4\upmu$L of the cultures into
$500\upmu$L of fresh media (1$\times$M9 with the carbon source)  every 48h for a total of
12 transfers at a dilution factor of 1:100,
roughly equivalent to 80 generations per culture (Fig. \mr).
Cycloheximide was not added to the media after the first two transfers.

@@

For each community, the most abundant colony morphotype at the end of the ninth transfer
was selected, resuspended in $100\upmu$L 1$\times$PBS and serially diluted (1:10).
Next, $20\upmu$L of the cells diluted to $10^{-6}$ were plated in the corresponding synthetic
minimal media and allowed to regrow at $30^\circ$C for 48h. Dominants were then inoculated
into $500\upmu$L of fresh media and incubated still at $30^\circ$C for 48h.
After this period, the communities stabilized for eleven transfers and the isolated dominants
were ready for the competition experiments (Fig \mr) at the onset of the twelfth transfer.

@@

All possible pairwise dominant-dominant and community-community
competition experiments
were performed by mixing equal volumes ($4\upmu$L) of each of the eight
communities or eight dominants at the onset of the twelfth transfer.
Competitions were set up in their native media,
i.e. in $500\upmu$L of 1$\times$M9 supplemented with 0.07 C-mol/L of
either glutamine or citrate in 96 deep-well plates.
Plates were incubated at $30^\circ$C for 48h.
Pairwise competitions were further propagated for seven serial transfers
(roughly 42 generations; Fig. \mr) by transferring $8\upmu$L of
each culture to fresh media ($500\upmu$L).

@@

The sequencing protocol was identical to that described in \cite{Goldford2018}.
Community samples were collected by spinning down at 3500rpm for 25min
in a bench-top centrifuge at room temperature;
cell pellets were stored at $-80^\circ$C before processing.
To maximize Gram-positive bacteria cell wall lysis,
the cell pellets were re-suspended and incubated at $37^\circ$C for 30min
in enzymatic lysis buffer (20mM Tris-HCl, 2mM sodium EDTA, 1.2\% Triton X-100)
and 20mg/mL of lysozyme from chicken egg white (Sigma-Aldrich).
After cell lysis, the DNA extraction and purification was performed using the
DNeasy 96 protocol for animal tissues (Qiagen).
The clean DNA in $100\upmu$L elution buffer of 10mM Tris-HCl, 0.5mM EDTA
at pH 9.0 was quantified using Quan-iT PicoGreen dsDNA Assay Kit
(Molecular Probes, Inc.)
and normalized to 5ng/$\upmu$L in nuclease-free water (Qiagen)
for subsequent 16S rRNA illumina sequencing.
16S rRNA amplicon library preparation was performed following a dual-index
paired-end approach \cite{Kozich2013}.
Briefly, PCR amplicon libraries of V4 regions of the 16S rRNA were prepared 
sing dual-index primers (F515/R805), then pooled and sequenced
using the Illumina MiSeq chemistry and platform.
Each sample went through a 30-cycle PCR in duplicate of $20\upmu$L
reaction volumes using 5ng of DNA each, dual index primers, and AccuPrime Pfx SuperMix (Invitrogen).
The thermocycling procedure includes a 2min initial denaturation step at
$95^\circ$C, and 30 cycles of the following PCR scheme:
(a) 20 second denaturation at $95^\circ$C,
(b) 15 second annealing at $55^\circ$C,
and (c) 5 minute extension at $72^\circ$C.
The duplicate PCR products of each sample were pooled, purified, and normalized
using SequalPrep PCR cleanup and normalization kit (Invitrogen).
Barcoded amplicon libraries were then pooled and sequenced using
Illumina Miseq v2 reagent kit, which generated 2$\times$250bp paired-end reads
at the Yale Center for Genome Analysis (YCGA).
The sequencing reads were demultiplexed on QIIME 1.9.0 \cite{Caporaso2010}.
The barcodes, indexes, and primers were removed from raw reads,
producing FASTQ files with both the forward and reverse reads for each sample,
ready for DADA2 analysis \cite{Callahan2017}.
DADA2 version 1.1.6 was used to infer unique biological exact sequence variants
(ESVs) for each sample
and na{\"i}ve Bayes was used to assign taxonomy using the SILVA version 123
database \cite{Wang2007,Quast2013}.
